Reproducible Research: Week 2 Project 
=======================================

## Loading and preprocessing the data

```{r} 
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip", mode="wb") 
unzip("activity.zip") 
stepdata <- read.csv("activity.csv", header = TRUE) 
head(stepdata)
```

## I. What is mean total number of steps taken per day?
### 1. Calculate the total number of steps taken per day
````{r}
StepsPerDay <- aggregate(stepdata$steps, list(stepdata$date), FUN=sum)
colnames(StepsPerDay) <- c("Date", "Steps")
head(StepsPerDay)
```

### 2.Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
g <- ggplot(StepsPerDay, aes(Steps))
g+geom_histogram(boundary=0, binwidth=2500, col="darkblue", fill="lightblue")+ggtitle("Histogram of steps per day")+xlab("Daily steps")+ylab("Frequency")+theme(plot.title = element_text(face="bold", size=12))
```

### 3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(StepsPerDay$Steps, na.rm=TRUE)
```

```{r}
median(StepsPerDay$Steps, na.rm=TRUE)
`````

## II. What is the average daily activity pattern?

### 1. A time series plot
Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
StepsPerTime <- aggregate(steps~interval,data=stepdata,FUN=mean,na.action=na.omit)
StepsPerTime$time <- StepsPerTime$interval/100
h <- ggplot(StepsPerTime, aes(time, steps))
h+geom_line(col="green")+ggtitle("Steps per time interval (average)")+xlab("Time")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=12))
```

### 2. 5-minute intervalwith the maximum number of steps
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
library(dplyr)
my_tbl_1 <- tibble::as_tibble(StepsPerTime)
my_tbl_1 %>% select(time, steps) %>% filter(steps==max(my_tbl_1$steps))
```

## III. Imputing missing values

### 1. Calculate and report the total number of missing values in the dataset
```{r}
my_tbl_2 <- tibble::as_tibble(stepdata)
my_tbl_2 %>% filter(is.na(steps)) %>% summarize(missing_values = n())
```

### 2. Filling in all of the missing values in the dataset

```{r}
stepdata$CompleteSteps <- ifelse(is.na(stepdata$steps), round(StepsPerTime$steps[match(stepdata$interval, StepsPerTime$interval)],0), stepdata$steps)
head(stepdata)
```

### 3. Create a new dataset with the missing data filled in

```{r}
activity_full <- data.frame(steps=stepdata$CompleteSteps, interval=stepdata$interval, date=stepdata$date)
head(activity_full)
```

### 4. Make a histogram of the total number of steps taken each day

```{r}
StepsPerDayFull <- aggregate(activity_full$steps, list(activity_full$date), FUN=sum)
colnames(StepsPerDayFull) <- c("Date", "Steps")
g <- ggplot(StepsPerDayFull, aes(Steps))
g+geom_histogram(boundary=0, binwidth=2500, col="darkgreen", fill="lightgreen")+ggtitle("Histogram of steps per day")+xlab("Steps")+ylab("Frequency")+theme(plot.title = element_text(face="bold", size=12))
```

### 5. Calculate and report the mean and median 

```{r}
mean(StepsPerDayFull$Steps)
```

```{r}
median(StepsPerDayFull$Steps)
```

Imputing missing data has only a little impact on the mean ant the median of the total daily number of steps. 

## IV. Are there differences in activity patterns between weekdays and weekends?

### 1. Create a new factor variable in the dataset with two levels - “weekday” and “weekend” indicating whether a given date is a weekday or weekend day

```{r}
activity_full$RealDate <- as.Date(activity_full$date, format = "%Y-%m-%d")
activity_full$weekday <- weekdays(activity_full$RealDate)
activity_full$DayType <- ifelse(activity_full$weekday=='Saturday' | activity_full$weekday=='Sunday', 'weekend','weekday')
head(activity_full, n=10)
```

### 2. Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
StepsPerTimeDT <- aggregate(steps~interval+DayType,data=activity_full, FUN=mean,na.action=na.omit)
StepsPerTimeDT$time <- StepsPerTime$interval/100
j <- ggplot(StepsPerTimeDT, aes(time, steps))
j+geom_line(col="red")+ggtitle("Average steps per time interval: weekdays vs. weekends")+xlab("Time")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=12))+facet_grid(DayType ~ .)
```





